---
title: "3Task"
author: "TomHollinberger"
date: "9/2/2020"
output: 
  html_document: 
    keep_md: yes
    toc: TRUE
    
---
```{r}

```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# ***Section 1:  DPLYR Practice***
```{r}
library(tidyverse)
library(dplyr)
iris
```



###  1. Arrange the iris data by Sepal.Length and display the first six rows.

```{r}
iris
sorted <- arrange(iris, Sepal.Length)
head(sorted)
```

###  2. Select the Species and Petal.Width columns and put them into a new data set called testdat.

```{r}
testdat <- select(iris,Species,Petal.Width)
testdat
```

###  3. Create a new table that has the mean for each variable by Species.
info on tables from https://www.cyclismo.org/tutorial/R/types.html
use the pipe from 5.6.1
This is like example in 5.6.1, 
but this gets error message, just like what is in the textbook 5.6.1

```{r}
byspecies <- (iris) %>%
  group_by(Species) %>%
  summarise(
    meanSL = mean(Sepal.Length, na.rm = TRUE),
    meanSW = mean(Sepal.Width, na.rm = TRUE),
    meanPL = mean(Petal.Length, na.rm = TRUE),
    meanPW = mean(Petal.Width, na.rm = TRUE)
  )
```
  
This works but doesn't create table
```{r}
iris %>% 
  group_by(Species) %>% 
  summarise(meanSL = mean(Sepal.Length, na.rm = TRUE),
            meanSW = mean(Sepal.Width, na.rm = TRUE),
            meanPL = mean(Petal.Length, na.rm = TRUE),
            meanPW = mean(Petal.Width, na.rm = TRUE))
```

###  4. Read about the ?summarize_all() function and get a new table with the means and standard deviations for each species.
?summarize_all()
```{r}
iris %>% 
  group_by(Species) %>% 
  summarise(meanSL = mean(Sepal.Length, na.rm = TRUE),
            meanSW = mean(Sepal.Width, na.rm = TRUE),
            meanPL = mean(Petal.Length, na.rm = TRUE),
            meanPW = mean(Petal.Width, na.rm = TRUE),
            stdevSL = sd(Sepal.Length, na.rm = TRUE),
            stdevSW = sd(Sepal.Width, na.rm = TRUE),
            stdevPL = sd(Petal.Length, na.rm = TRUE),
            stdevPW = sd(Petal.Width, na.rm = TRUE))
```



If you want to apply multiple transformations, pass a list of
functions. When there are multiple functions, they create new
variables instead of modifying the variables in place:
This works in Rscript, but doesn't create a table, and has clunky variable names.
But gets an error when knitting from RMD. So I disabled it single quotes.

```{r}
'by_species %>%
  summarise_all(list(mean,sd))'
```






# ***Section 2  :  Novel Questions about data***
* What are the predictors of academic success a WJC?
* What define academic success?  GPA, Course level of difficulty, Major level of difficulty, Graduation YEs or No?
* What data is available?
* What vintage is the predictor data (high school info, college freshman indicators)
* What vintage is the result data (graduation within the last # years)
* Is high school data in a consistent format across different high schools?
* Talked at length with Dr Baker and Ms Gladbach about this.  
* They have lot of the same questions.  And would use the answers to target marketing to prospective students whose profile indicates would be successful at WJC.






























