---
title: "3CaseStudy"
author: "TomHollinberger"
date: "9/7/2020"
output: 
  html_document: 
    keep_md: yes
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(tidyverse)  #has function to read files
df_orig <- read_csv("E:/000 DTS 350 Data Visualization/DTS350-hollinbergert/DTS350TemplateMaster/Week_03/3CaseStudy/CensusAtSchool.csv")

str(df_orig)
df_orig
df_inch <- df_orig
df_inch <- mutate(df_inch,Height = Height/2.54, Foot_Length = Foot_Length/2.54, Arm_Span = Arm_Span/2.54 )
head(df_inch, n=10)
df_environment <- filter(df_inch, Importance_reducing_pollution >= 750 | Importance_recycling_rubbish >=750 |	Importance_conserving_water	>=750 | Importance_saving_enery >=750 | Importance_owning_computer >=750 |	Importance_Internet_access >=750)
#would be nice to "filter if not start with Import"
View(df_environment)
df_environment <- (arrange(df_environment, Ageyears))
df_environment
head(df_environment, n=10)
df_extra <- select(df_orig, Country,	Region,	Gender,	Ageyears,	Handed,	Height,	Foot_Length,	Arm_Span,	Languages_spoken,	Travel_to_School,	Travel_time_to_School,	Reaction_time,	Score_in_memory_game,	Favourite_physical_activity)
#would be nice to "select if not start with"
#easy to copy and paste column headers from excel, then insert commas
head(df_extra, n=10)
df_numbers <- df_orig
df_numbers <- group_by(df_numbers,Country,Gender)
df_numbers <- summarise(df_numbers, count = n())
head(df_numbers, n=10)
df_orig
df_CountryGenderImportances <- select(df_orig, Country, Gender, Importance_reducing_pollution, Importance_recycling_rubbish, Importance_conserving_water, Importance_saving_enery,	Importance_owning_computer,	Importance_Internet_access)
head(df_CountryGenderImportances, n=10)

df_CountryGenderGrp <- group_by(df_CountryGenderImportances, Country, Gender)
df_CountryGenderGrp
df_CountryGenderGrpSum <- summarise(df_CountryGenderGrp)
View(df_CountryGenderGrp)
df_CountryGenderGrpSum <- summarise(df_CountryGenderGrp,
                       Count_reducing_pollution = sum(!is.na(Importance_reducing_pollution)), 
                       Mean_reducing_pollution = mean(Importance_reducing_pollution, na.rm = TRUE),
                       Standdev_reducing_pollution = sd(Importance_reducing_pollution, na.rm = TRUE),
                       Count_recycling_rubbish = sum(!is.na(Importance_recycling_rubbish)),                     
                       Mean_recycling_rubbish = mean(Importance_recycling_rubbish, na.rm = TRUE),
                       Standdev_recycling_rubbish = sd(Importance_recycling_rubbish, na.rm = TRUE),
                       Count_conserving_water = sum(!is.na(Importance_conserving_water)),
                       Mean_conserving_water = mean(Importance_conserving_water, na.rm = TRUE),
                       Standdev_conserving_water = sd(Importance_conserving_water, na.rm = TRUE),
                       Count_saving_enery = sum(!is.na(Importance_saving_enery)),
                       Mean_saving_enery = mean(Importance_saving_enery, na.rm = TRUE),
                       Standdev_saving_enery = sd(Importance_saving_enery, na.rm = TRUE),
                       Count_owning_computer = sum(!is.na(Importance_owning_computer)),
                       Mean_owning_computer = mean(Importance_owning_computer, na.rm = TRUE),
                       Standdev_owning_computer = sd(Importance_owning_computer, na.rm = TRUE),
                       Count_Internet_access = sum(!is.na(Importance_Internet_access)),
                       Mean_Internet_access = mean(Importance_Internet_access, na.rm = TRUE),
                       Standdev_Internet_access = sd(Importance_Internet_access, na.rm = TRUE))
df_CountryGenderGrpSum
View(df_CountryGenderGrpSum)
```


## [8 ] Create an .Rmd file that has two sections:
###  Section 1: Give the first 10 rows of each of the data frames you created
```{r}
head(df_orig, n = 10)
head(df_inch, n = 10)
head(df_environment, n = 10)
head(df_extra, n = 10)
head(df_numbers, n = 10)
head(df_CountryGenderImportances, n = 10)
head(df_CountryGenderGrp, n = 10)
head(df_CountryGenderGrpSum, n = 10)

```







### Section 2: Include a summary of your findings to the question which includes at least one visualization to support your findings.

```{r}
ggplot(data = df_CountryGenderGrp) +
geom_boxplot(mapping = aes(x = Gender, y = Importance_reducing_pollution)) +
  facet_wrap(~ Country, nrow = 1)

ggplot(data = df_CountryGenderGrp) +
  geom_boxplot(mapping = aes(x = Gender, y = Importance_recycling_rubbish)) +
  facet_wrap(~ Country, nrow = 1)

ggplot(data = df_CountryGenderGrp) +
  geom_boxplot(mapping = aes(x = Gender, y = Importance_conserving_water)) + 
  facet_wrap(~ Country, nrow = 1)

ggplot(data = df_CountryGenderGrp) +    
  geom_boxplot(mapping = aes(x = Gender, y = Importance_saving_enery)) +
  facet_wrap(~ Country, nrow = 1)

ggplot(data = df_CountryGenderGrp) +
  geom_boxplot(mapping = aes(x = Gender, y = Importance_owning_computer)) + 
  facet_wrap(~ Country, nrow = 1)

ggplot(data = df_CountryGenderGrp) +  
  geom_boxplot(mapping = aes(x = Gender, y = Importance_Internet_access)) +
  facet_wrap(~ Country, nrow = 1)

#How to stitch these together in one chart?  So that distinct variables are the groups(data series)
```


#### **Summary Observations:**

1.  Males, across the board, and especially in USA, gave lower scores, and are less concerned about environmental matters.
2.  Males are generally more concerned (gave higher scores) to computer-related issues.
     
